---
title: "markdown"
author: "Vincent Blain"
date: "June 10, 2016"
output: html_document
---

## possible packages needed

```{r demo, eval=FALSE}
install.packages("R.utils")
library(R.utils)
install.packages("data.table")
library(data.table)
install.packages("dplyr")
library(dplyr)
install.packages("quantmod")
library(quantmod)
install.packages("stringr")
library(stringr)
install.packages("qdap")
library(qdap)
install.packages("plyr")
library(plyr)
library(knitr)
```

## 1. Merges the training and the test sets to create one data set.

Download File

```{r, eval=TRUE}
url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
f <- file.path(getwd(), "getdata_projectfiles_UCI_HAR_Dataset.zip")
download.file(url, f)
```

unzip file install R.utils package to get to unzip command and load the library before using

```{r, eval=TRUE}
unzip(f, overwrite = TRUE)
fPath <- file.path(getwd(), "UCI HAR Dataset")
files <- list.files(fPath, recursive = TRUE)
```
output list of files
```{r, echo=TRUE}
files
```

ignore "inertial signals"
```{r, eval=TRUE}
filesFinal <- files[ !grepl("Inertial Signals",files) ]
```

output list of files without "inertial signals"
```{r, echo=TRUE}
filesFinal
```

load data
```{r, eval=TRUE}
d_Test_Subject   <- read.table(file.path(fPath, "test" , "subject_test.txt"), header = FALSE)
d_Test_Activity  <- read.table(file.path(fPath, "test" , "Y_test.txt" ), header = FALSE)
d_Test_Features  <- read.table(file.path(fPath, "test" , "X_test.txt" ), header = FALSE)
```
output "d_Test_Subject"
```{r, echo=TRUE}
str(d_Test_Subject)
```
output "d_Test_Activity"
```{r, echo=TRUE}
str(d_Test_Activity)
```
output "d_Test_Features"
```{r, echo=TRUE}
str(d_Test_Features)
```

```{r, eval=TRUE}
d_Train_Subject  <- read.table(file.path(fPath, "train", "subject_train.txt"), header = FALSE)
d_Train_Activity <- read.table(file.path(fPath, "train", "Y_train.txt"), header = FALSE)
d_Train_Features <- read.table(file.path(fPath, "train", "X_train.txt"), header = FALSE)
```
output "d_Train_Subject"
```{r, echo=TRUE}
str(d_Train_Subject)
```
output "d_Train_Activity"
```{r, echo=TRUE}
str(d_Train_Activity)
```
output "d_Train_Features"
```{r, echo=TRUE}
str(d_Train_Features)
```

merge datasets
```{r, eval=TRUE}
dfSubject  <- rbind(d_Train_Subject, d_Test_Subject)
dfActivity <- rbind(d_Train_Activity, d_Test_Activity)
dfFeatures <- rbind(d_Train_Features, d_Test_Features)
```

set names to vars
```{r, eval=TRUE}
names(dfSubject)  <-c("subject")
names(dfActivity) <- c("activity")
```

load features names remove bad characters
```{r, eval=TRUE}
FeaturesNames     <- read.table(file.path(fPath, "features.txt"),head=FALSE)
FeaturesNames$V2 <- gsub("(", "", fixed = TRUE, FeaturesNames$V2)
FeaturesNames$V2 <- gsub(")", "", fixed = TRUE, FeaturesNames$V2)
FeaturesNames$V2 <- gsub("-", "", fixed = TRUE, FeaturesNames$V2)
FeaturesNames$V2 <- gsub(",", "", fixed = TRUE, FeaturesNames$V2)
FeaturesNames$V2 <- gsub(" ", "", fixed = TRUE, FeaturesNames$V2)
```

output "FeaturesNames$V2"
```{r, echo=TRUE}
FeaturesNames$V2
```

 make columns lower case
```{r, eval=TRUE}
names(dfFeatures) <- tolower(FeaturesNames$V2)
```
output "FeaturesNames$V2"
```{r, echo=TRUE}
str(dfFeatures)
```

merge datasets
```{r, eval=TRUE}
dfCombine <- cbind(dfSubject, dfActivity)
dfFinal   <- cbind(dfCombine, dfFeatures)
```

output "dfFinal" with updates
```{r, echo=TRUE}
str(dfFinal)
```

## 2. Extracts only the measurements on the mean and standard deviation for each measurement.

get only the columns that match mean or std, make sure to set to lower case
```{r, eval=TRUE}
SubFeats <- tolower(FeaturesNames$V2[grep("mean|std", FeaturesNames$V2)])
```
add the subject and activity columns also
```{r, eval=TRUE}
setColNames <- c(as.character(SubFeats), "subject", "activity" )
```
get the subset of data based on selected columns
```{r, eval=TRUE}
almostTidyData <- subset(dfFinal, select=setColNames)
```

output "almostTidyData" should contain means, std, subject, activity
```{r, echo=TRUE}
str(almostTidyData)
```

## 3. Uses descriptive activity names to name the activities in the data set
```{r, eval=TRUE}
ativities <- read.table(file.path(fPath, "activity_labels.txt"),header = FALSE)
```

output "ativities"
```{r, echo=TRUE}
ativities
```

replace the activities integer values with their corresponding string values
```{r, eval=TRUE}
almostTidyData$activity<-factor(almostTidyData$activity);
almostTidyData$activity<- factor(almostTidyData$activity,labels=as.character(ativities$V2))
```

output "almostTidyData"
```{r, echo=TRUE}
head(almostTidyData$activity,10)
```

## 4. Appropriately labels the data set with descriptive variable names.
```{r, eval=TRUE}
names(almostTidyData)<-gsub("^t", "time", names(almostTidyData))
names(almostTidyData)<-gsub("^f", "frequency", names(almostTidyData))
names(almostTidyData)<-gsub("acc", "accelerometer", names(almostTidyData))
names(almostTidyData)<-gsub("gyro", "gyroscope", names(almostTidyData))
names(almostTidyData)<-gsub("mag", "magnitude", names(almostTidyData))
names(almostTidyData)<-gsub("bodybody", "body", names(almostTidyData))
```

output "almostTidyData" with tidy column names
```{r, echo=TRUE}
names(almostTidyData)
```

## 5. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.
```{r, eval=TRUE}
tidyDf<-aggregate(. ~subject + activity, almostTidyData, mean)
tidyDf<-tidyDf[order(tidyDf$subject,tidyDf$activity),]
write.table(tidyDf, file = "tidydata.txt",row.name=FALSE)
```
output "almostTidyData" with tidy column names (first 6 columns only)
```{r, echo=TRUE}
head(tidyDf[1:6], 10)
```
